---
title: "Time Series Regression: Part 2"
subtitle: "PLSC 505"
author: "Professor Linn"
date: "February 22, 2022"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '3'
    toc_float: true
    code_folding: hide
    mathjax: "https://cdnjs.cloudflare.com/ajax/libs/mathjax/2.7.5/MathJax.js?config=TeX-AMS_CHTML.js"
  html_notebook:
    number_sections: no
    theme: cerulean
    toc: yes
    toc_depth: 3
  pdf_document:
    toc: yes
    toc_depth: '3'
  word_document: null
---
<style type="text/css">

body{ /* Normal  */
      font-size: 14px;
  }
td {  /* Table  */
  font-size: 18px;
}

code.r{ /* Code block */
    font-size: 14px;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 14px;
}
.nobullet li {
  list-style-type: none;
}
}
</style>



# Preliminaries

Load packages and read in the data:

```{r, message=FALSE, warning=FALSE}
library(dynlm)  # To run dynamic LS
library(stargazer) # For table output
library(msm) # for delta method
library(lmtest) # for LM tests
library(sandwich) # for various variance-covariance matrices
library(dplyr) # to manipulate data frames
library(ggplot2)
```

```{r}
load("TSRegressionApproval.RData")
```


```{r}
AppEcon <- AppEcon %>%
  filter(year >= 1978 & year <= 2020)
```

```{r}
AppEcon.ts <- ts(AppEcon, start = c(1978,1), end = c(2020,12), freq = 12)
```


# Step Three: Simplify the General Model


## Can Higher Order Lags be Dropped from the Model?

### F-tests

+ Start with ADL(1,3;3) model of approval, which has one more lag than the model we considered last week. I'll call it ADL.1.3.3.
+ Ewtimate the ADL(1,2;3) model from last week.
+ Test whether we can restrict all the lag 3 coefficients to zero: $H_0: \beta^{DSPIC96_pc_pca}_3=\beta^{lei_pc}_3=\beta^{cei_pc}_3=0$.
+ Note that we need to make sure both models are estimated over the same time period. I'm going to use the start and stop arguments in `dynlm()` to fix the sample period to be the same and estimate the ADL(1,3;3) model.


```{r, echoe=FALSE}
#General Model
ADL.1.3.3 <- dynlm(Approving~L(Approving,1) +  
                     DSPIC96_pc_pca +L(DSPIC96_pc_pca, c(1,2,3))+ 
                     lei_pc + L(lei_pc, c(1,2,3)) + 
                     cei_pc + L(cei_pc, c(1,2,3)) + 
                     Reagan + BushI + Clinton + BushII + Obama + 
                     Trump  + Honey + Inaug +  RAss + IraqKuwait + 
                     Desert + Lehman + Sept11 + Mueller  + Impeach1 +
                     Impeach2 + CovidP + CovidS + ObamaJan + BushJan +
                     hostages, start=c(1978,4), end=c(2020,12), data=AppEcon.ts)
```

```{r, message=FALSE, warning=FALSE}
#Drop Lag 3
ADL.1.2.3 <- dynlm(Approving~L(Approving,1) +  
                     DSPIC96_pc_pca +L(DSPIC96_pc_pca, c(1,2))+ 
                     lei_pc + L(lei_pc, c(1,2)) + 
                     cei_pc + L(cei_pc, c(1,2)) + 
                     Reagan + BushI + Clinton + BushII + Obama + 
                     Trump  + Honey + Inaug +  RAss + IraqKuwait + 
                     Desert + Lehman + Sept11 + Mueller  + Impeach1 +
                     Impeach2 + CovidP + CovidS + ObamaJan + BushJan +
                     hostages, start=c(1978,4), end=c(2020,12), data=AppEcon.ts)
```

+ Use the `anova()` function to conduct the $F$-test. This function takes the unrestricted model (ADL.1.3.3.) and the restricted model (ADL.1.2.3.) as arguments. It conducts an F-test to determine whether the restriction is valid. The null hypothesis is that the restriction is binding.


```{r}
anova(ADL.1.3.3, ADL.1.2.3)
```


Based on the F-test results, we **cannot** reject the null hypothesis that $\beta^{DSPIC96_pc_pca}_3=\beta^{lei_pc}_3=\beta^{cei_pc}_3=0$.

+ Make sure the restricted model is dynamically complete (which we did last week).

```{r}
bgtest(ADL.1.2.3, order=24, type="Chisq")
bgtest(ADL.1.2.3, order=12, type="Chisq")
bgtest(ADL.1.2.3, order=6, type="Chisq")
```



Now let's test whether we can drop lag 2 of these regressors: $H_0: \beta^{DSPIC96_pc_pca}_2=\beta^{lei_pc}_2=\beta^{cei_pc}_2=0$.

```{r, message=FALSE, warning=FALSE}
#General model
ADL.1.2.3 <- dynlm(Approving~L(Approving,1) +  
                     DSPIC96_pc_pca +L(DSPIC96_pc_pca, c(1,2))+ 
                     lei_pc + L(lei_pc, c(1,2)) + 
                     cei_pc + L(cei_pc, c(1,2)) + 
                     Reagan + BushI + Clinton + BushII + Obama + 
                     Trump  + Honey + Inaug +  RAss + IraqKuwait + 
                     Desert + Lehman + Sept11 + Mueller  + Impeach1 +
                     Impeach2 + CovidP + CovidS + ObamaJan + BushJan  + hostages, start=c(1978,3), end=c(2020,12), data=AppEcon.ts)
```




```{r, echoe=FALSE}
#Drop L2
ADL.1.1.3 <- dynlm(Approving~L(Approving,1) +  
                     DSPIC96_pc_pca +L(DSPIC96_pc_pca, 1)+ 
                     lei_pc + L(lei_pc, 1) + 
                     cei_pc + L(cei_pc, 1) + 
                     Reagan + BushI + Clinton + BushII + Obama + 
                     Trump  + Honey + Inaug +  RAss + IraqKuwait + 
                     Desert + Lehman + Sept11 + Mueller  + Impeach1 +
                     Impeach2 + CovidP + CovidS + ObamaJan + BushJan +
                     hostages, start=c(1978,3), end=c(2020,12), data=AppEcon.ts)
```

```{r}
#Test if this restriction is valid
anova(ADL.1.2.3, ADL.1.1.3)

```


Based on the F-test results, we **reject** the null hypothesis that $\beta^{DSPIC96_pc_pca}_2=\beta^{lei_pc}_2=\beta^{cei_pc}_2=0$.

**If** we had failed to reject, we would need to check whether the restricted model was dynamically complete by testing for serial correlation in the model residuals using the Breusch-Godfrey LM test. If it was not, we would not accept the restriction.



**We could have started with the GECM and tested whether the highest lagged differences are jointly zero.**

### Information Criteria

Compare the AIC and BIC for our three models.

We need all models estimated on same sample!


```{r}
ADL.1.2.3.S <- dynlm(Approving~L(Approving,1) +  
                     DSPIC96_pc_pca +L(DSPIC96_pc_pca, c(1,2))+ 
                     lei_pc + L(lei_pc, c(1,2)) + 
                     cei_pc + L(cei_pc, c(1,2)) + 
                     Reagan + BushI + Clinton + BushII + Obama + 
                     Trump  + Honey + Inaug +  RAss + IraqKuwait + 
                     Desert + Lehman + Sept11 + Mueller  + Impeach1 +
                     Impeach2 + CovidP + CovidS + ObamaJan + BushJan  + hostages, start=c(1978,4), end=c(2020,12), data=AppEcon.ts)
ADL.1.1.3.S <- dynlm(Approving~L(Approving,1) +  
                     DSPIC96_pc_pca +L(DSPIC96_pc_pca, 1)+ 
                     lei_pc + L(lei_pc, 1) + 
                     cei_pc + L(cei_pc, 1) + 
                     Reagan + BushI + Clinton + BushII + Obama + 
                     Trump  + Honey + Inaug +  RAss + IraqKuwait + 
                     Desert + Lehman + Sept11 + Mueller  + Impeach1 +
                     Impeach2 + CovidP + CovidS + ObamaJan + BushJan +
                     hostages, start=c(1978,4), end=c(2020,12), data=AppEcon.ts)

```


```{r}
InfoAIC <- rbind(AIC(ADL.1.3.3),AIC(ADL.1.2.3.S), AIC(ADL.1.1.3.S))
InfoBIC <- rbind(BIC(ADL.1.3.3), BIC(ADL.1.2.3.S), BIC(ADL.1.1.3.S))
ModelName <- rbind("ADL(1,3;3)","ADL(1,2;3)","ADL(1,1;3)")
InfoCr <- data.frame(ModelName, InfoAIC, InfoBIC)
InfoCr
```


The AIC and BIC select the ADL(1,2;3) model, consistent with the F-test results.

# Step Four (If this was our final model)


```{r}
stargazer(ADL.1.2.3, type="text",
          omit=c("Reagan", "BushI", "Clinton", "BushII", "Obama", "Trump",
                 "Honey", "Inaug", "RAss", "IraqKuwait", "Desert",
                 "Lehman", "Sept11", "Mueller", "Impeach1", "Impeach2",
                 "CovidP", "CovidS", "ObamaJan", "BushJan", "hostages"))

```

## Impact Multipliers

Let's calculate the direct effect of a standard deviation change in $x$ for our model. We multiple the size of the change in $x$ by the individual impact multipliers.

```{r}
print(paste("St Dev DI = ", round(sd(AppEcon$DSPIC96_pc_pca),3)))
print(paste("Lag 0 Direct Effect DI = ", round(ADL.1.2.3$coefficients[3]*sd(AppEcon$DSPIC96_pc_pca), 3)))
print(paste("Lag 1 Direct Effect DI= ", round(ADL.1.2.3$coefficients[4]*sd(AppEcon$DSPIC96_pc_pca), 3)))
print(paste("Lag 2 Direct Effect DI = ", round(ADL.1.2.3$coefficients[5]*sd(AppEcon$DSPIC96_pc_pca), 3)))
```

Approval is expected to increase by about 0.29 at time $t$, by 0.51 at $t+1$, and 0.38 at $t+2$ following a standard deviation change in disposable income growth.


Do the same for LEI and CEI growth rates.


## The Total Effect: The Long Run Multiplier (LRM)

+ The LRM for disposable income growth is given by approximately <span style="color:purple">$\frac{0.28 + 0.51 + 0.37}{1-0.9}=11.6$</span>. The total effect of a standard deviation change in disposable income growth can be calculated as follows:

```{r}
LRM.DI <- (ADL.1.2.3$coefficients[3] + ADL.1.2.3$coefficients[4] + ADL.1.2.3$coefficients[5])/(1- ADL.1.2.3$coefficients[2])
LRM.DI
LRM.DI*sd(AppEcon$DSPIC96_pc_pca)
```


Do the same for growth in LEI and CEI:



### Computing the Standard Error of the LRM: The Delta Method. 


Use the `deltamethod()` function in the msm package.

+ The first argument is the formula for the LRM of interest, e.g., here x7 is the coefficient on contemporaneous consumer sentiment. 
+ The next argument specifies  the model fit object containing the coefficients of interest: coef(LM3.adl11).
+ The last argument specifies which error variance-covariance matrix to use. Since we have heteroscedasticity, we specify cov=vcovHC(ADL.1.2.3, type="HC1"). (Note: We need to load the sandwich package to specify heteroscedastic robust variance-covariance matrix.)

```{r}
LRM.DI.se <- deltamethod(~(x3+x4+x5)/(1-x2), coef(ADL.1.2.3), cov=vcovHC(ADL.1.2.3, type="HC1"))
print(paste("LRM DI = ", round(LRM.DI,2)))
print(paste("LRM DI se = ", round(LRM.DI.se,2)))
```

The LRM for disposable income growth is significantly different from zero at .05 level ($|LRM/LRM_{se}|>1.96$).

Test the significance of the LEI and CEI growth LRM



## The Long-Run Equilibirum


We can calculate the long-run equilibrium value of presidential approval (ignoring the interventions)


\begin{equation*}
A^{*}=-3.40 + frac{0.27 + 0.51 + 0.38}{1-0.9}DI^*  +  \frac{9.7-21-22.7}{1-0.9}LEI^* + \frac{(23.99+ 40.19+121.38)}{(1-0.9)}CEI^*.
\end{equation*}

If we calculate the mean of disposable income growth, the leading economic indicator index, and the coincident economic indicator index, we can plug and chug.

```{r}
means <- AppEcon %>%
  filter(date>"1978/3/1"&date<"2020/12/1") %>%
  summarise(DI.bar= mean(DSPIC96_pc_pca), LEI.bar=mean(lei_pc), CEI.bar=mean(cei_pc))
means
```


```{r}
App.Equil <- (ADL.1.2.3$coefficients[1] + 
  (ADL.1.2.3$coefficients[3] + ADL.1.2.3$coefficients[4]+ ADL.1.2.3$coefficients[5])*means$DI.bar  +
  (ADL.1.2.3$coefficients[6]+ADL.1.2.3$coefficients[7]+ ADL.1.2.3$coefficients[8])*means$LEI.bar + 
  (ADL.1.2.3$coefficients[9]+ADL.1.2.3$coefficients[10]+ ADL.1.2.3$coefficients[11])*means$CEI.bar)/(1- ADL.1.2.3$coefficients[2])
App.Equil
```

The long-run equilibrium value of presidential approval over this time period is about 41.7%.



## How Does the Effect of a Change in an Independent Variable Play Out Over Time?

### Period-to-Period Effects: The (Simple) Impulse Response Function

Calculate and plot the SIRF.

First, we need to create an empty data frame to hold the number of lags for which we want to calculate the irf (**period**) and the values of predicted approval given disposable income growth (**DIIRF**).

```{r}
DI.irf <- data.frame(period = c(0:36),
                           DIIRF = NA_real_)
```

Next we generate the predicted values of approval given a one unit shock to consumer sentiment. **We are doing the same thing we did in the slides.**

```{r}
#First period
DI.irf$DIIRF[1] <- ADL.1.2.3$coefficients[2]*0 +
                                    ADL.1.2.3$coefficients[3]*1 +
                                    ADL.1.2.3$coefficients[4]*0 +
                                    ADL.1.2.3$coefficients[5]*0
#Second period
DI.irf$DIIRF[2] <- ADL.1.2.3$coefficients[2]*DI.irf$DIIRF[1] +
                                    ADL.1.2.3$coefficients[3]*0 +
                                    ADL.1.2.3$coefficients[4]*1 +
                                    ADL.1.2.3$coefficients[5]*0
#Third period
DI.irf$DIIRF[3] <- ADL.1.2.3$coefficients[2]*DI.irf$DIIRF[2] +
                                    ADL.1.2.3$coefficients[3]*0 +
                                    ADL.1.2.3$coefficients[4]*0 +
                                    ADL.1.2.3$coefficients[5]*1

#Periods 4-36
for (t in 4:37) {
        DI.irf$DIIRF[t] <- ADL.1.2.3$coefficients[2]*DI.irf$DIIRF[t-1] +
                                    ADL.1.2.3$coefficients[3]*0 +
                                    ADL.1.2.3$coefficients[4]*0 +
                                    ADL.1.2.3$coefficients[5]*0
        }

```

Finally, we can plot the result.

```{r}
#### Plot
ggplot(data = DI.irf, aes(x = period,y = DIIRF)) +
               geom_bar(stat = "identity") +
               geom_hline(aes(yintercept = 0),
                          colour = "#990000",
                          linetype = "dashed") +
               ylab("IRF") +
               xlab("Period") +
               ggtitle("Impulse Response Function Plot (Disposable Income Growth)") +
               theme_bw()
```


IRF of growth in the Leading Indicator Index in the ADL(1,2;3) model.

```{r, echo=TRUE}
LEI.irf <- data.frame(period = c(0:36),
                           LEIIRF = NA_real_)

#First period
LEI.irf$LEIIRF[1] <- ADL.1.2.3$coefficients[2]*0 +
                                    ADL.1.2.3$coefficients[6]*1 +
                                    ADL.1.2.3$coefficients[7]*0 +
                                    ADL.1.2.3$coefficients[8]*0
#Second period
LEI.irf$LEIIRF[2] <- ADL.1.2.3$coefficients[2]*LEI.irf$LEIIRF[1] +
                                    ADL.1.2.3$coefficients[6]*0 +
                                    ADL.1.2.3$coefficients[7]*1 +
                                    ADL.1.2.3$coefficients[8]*0
#Third period
LEI.irf$LEIIRF[3] <- ADL.1.2.3$coefficients[2]*LEI.irf$LEIIRF[2] +
                                    ADL.1.2.3$coefficients[6]*0 +
                                    ADL.1.2.3$coefficients[7]*0 +
                                    ADL.1.2.3$coefficients[8]*1

#Periods 4-36
for (t in 4:37) {
        LEI.irf$LEIIRF[t] <- ADL.1.2.3$coefficients[2]*LEI.irf$LEIIRF[t-1] +
                                    ADL.1.2.3$coefficients[3]*0 +
                                    ADL.1.2.3$coefficients[4]*0 +
                                    ADL.1.2.3$coefficients[5]*0
        }


ggplot(data = LEI.irf, aes(x = period,y = LEIIRF)) +
               geom_bar(stat = "identity") +
               geom_hline(aes(yintercept = 0),
                          colour = "#990000",
                          linetype = "dashed") +
               ylab("IRF") +
               xlab("Period") +
               ggtitle("Impulse Response Function Plot (Growth in Leading Economic Indicator)") +
               theme_bw()
```

Create the IRF of Growth in Coincident Economic Indicator Index in the ADL(1,2;3) model.



### The Cumulative Impulse Response Function

We can see how the effects of the shock cumulate by adding up the individual period effects in the IRF to create the cumulative IRF or CIRF. The cumulative effect at time $t=0$ is the value of the IRF at $t=0$. At $t+1$ the effect is the sum of the time $t=0$ effect and the time $t=t+1$ effect or. The cumulative effect at time $t+2$ is the sum of the individual effects at $t=0$, $t=t+1$ and $t=t+2$. 

```{r}
DI.cirf <- data.frame(period = c(0:36),
                           cDIIRF = NA_real_)
```

```{r}
#First period
DI.cirf$cDIIRF[1] <- ADL.1.2.3$coefficients[2]*0 +
                              ADL.1.2.3$coefficients[3]*1 +
                              ADL.1.2.3$coefficients[4]*0 +
                              ADL.1.2.3$coefficients[5]*0
currentvalue <- DI.cirf$cDIIRF[1]
#Rest of periods
for (t in 2:37) {

    DI.cirf$cDIIRF[t] <- DI.irf$DIIRF[t] +
                                 currentvalue
    currentvalue <- DI.cirf$cDIIRF[t]
}
ggplot(data = DI.cirf, aes(x = period, y = cDIIRF)) +
                geom_bar(stat = "identity") +
                ylab("CIRF") +
                xlab("Period") +
                ggtitle("Cumulative Impulse Response Function Plot (Disposable Income Growth)") +
                theme_bw()
```

Notice that the cumulative effect is exactly the long run multiplier.


```{r, echo=TRUE}
LEI.cirf <- data.frame(period = c(0:36),
                           cLEIIRF = NA_real_)

#First period
LEI.cirf$cLEIIRF[1] <- ADL.1.2.3$coefficients[2]*0 +
                       ADL.1.2.3$coefficients[6]*1 +
                       ADL.1.2.3$coefficients[7]*0 +
                       ADL.1.2.3$coefficients[8]*0
currentvalue <- LEI.cirf$cLEIIRF[1]
#Rest of periods
for (t in 2:37) {

    LEI.cirf$cLEIIRF[t] <- LEI.irf$LEIIRF[t] + currentvalue
    currentvalue <- LEI.cirf$cLEIIRF[t]
}
ggplot(data = LEI.cirf, aes(x = period, y = cLEIIRF)) +
                geom_bar(stat = "identity") +
                ylab("CIRF") +
                xlab("Period") +
                ggtitle("Cumulative Impulse Response Function Plot (Growth in Leading Economic Indicator Index)") +
                theme_bw()
```

Create the cumulative IRF for CEI.

### The Mean and Median Lag Length


In order to determine the median and mean lag lengths we standardize the individual period cumulative effects as a proportion of the LRM. The LRM  growth in disposable income  in this example is given by $\frac{\sum_{i=1}^q\beta_{j,i}}{1-\sum_{i=1}^p\alpha_i}=\frac{0.27 + 0.51 + 0.38}{1-0.9}=11.6$.


Half of the effect is 0.5*11.6=5.8. By  listing the values of the CIRF, we can see when we've reached this mark. (Sometimes the plot will be clear.)

```{r}
DI.cirf$cDIIRF
```

The eighth value is the first value greater than 5.8, so the median lag length is eight.

Alternatively we can standardize to report the cumulative impulse responses as a fraction of the LRM:

```{r}
DI.cirf$cDIIRF/LRM.DI
```

Now we see that half the effect (51.796%) has played out by lag 8.

```{r}
which.max((DI.cirf$cDIIRF/LRM.DI)>=0.5)
```


The mean lag length is more subjective. Here we can see that it takes almost three years for 97% of the effect to play out.

```{r}
which.max((DI.cirf$cDIIRF/LRM.DI)>=0.97)
```


For the LEI growth rate we have:

```{r, echo=TRUE}
print(paste("Standardized CIRF Leading Economic Indicator Index = "))
LEI.cirf$cLEIIRF/LRM.LEI

print(paste("Median Lag = ",
            which.max(LEI.cirf$cLEIIRF/LRM.LEI>=0.5)  ))

print(paste("Mean Lag = ",
            which.max(LEI.cirf$cLEIIRF/LRM.LEI>=0.97)  ))
```

Calculate the median and mean lag for the CEI.



# What Have We Learned?


Variable | $\beta_0$|  $\beta_1$ | $\beta_2$ | LRM | LRM se | Median Lag | Mean Lag
--------- | ---------| ----------- | ----- | ------- |----------- | ------- | ------
$DI$ | 0.282 |  0.509 | 0.371 | 11.6 | 3.69 | 8 | 35
$LEI$ | 11.065 | -20.916| -24.105 | -338.84 | 369.55 | 9| 36
$CEI$ | 24.286 | 41.708| 121.129 | 1868.23 | 628.51 | 9 | 35
---

Error correction rate: -0.10.



# Step Three Continued



## Let's try to simplify our ADL(1,2;3) model.

```{r, message=FALSE, warning=FALSE, echo=TRUE}
stargazer(ADL.1.2.3, type="text",
          summary=TRUE,
          report="vcst*",
          keep.stat=c("aic","bic","n","rsq","adj.rsq","ll","F","ser"),
          omit=c("Reagan", "BushI", "Clinton", "BushII", "Obama", "Trump",
                 "Honey", "Inaug", "RAss", "IraqKuwait", "Desert",
                 "Lehman", "Sept11", "Mueller", "Impeach1", "Impeach2",
                 "CovidP", "CovidS", "ObamaJan", "BushJan", "hostages"))
```


### Drop lag 2 LEI?

```{r, echo=TRUE}
ADL.1.2.3.a <- dynlm(Approving~L(Approving,1) +  
                       DSPIC96_pc_pca +L(DSPIC96_pc_pca, c(1,2))+ 
                       lei_pc + L(lei_pc, c(1)) + 
                       cei_pc + L(cei_pc, c(1,2)) + 
                       Reagan + BushI + Clinton + BushII + Obama + Trump  +
                       Honey + Inaug +  RAss + IraqKuwait + Desert + Lehman + 
                       Sept11 + Mueller  + Impeach1 + Impeach2 + CovidP + CovidS + 
                       ObamaJan + BushJan  + hostages, start=c(1978,4), end=c(2020,12), data=AppEcon.ts)
bgtest(ADL.1.2.3.a, order=12, type="Chisq")
bgtest(ADL.1.2.3.a, order=6, type="Chisq")
#summary(LM3a)

newRow <- data.frame(ModelName='ADL.1.2.3.a',InfoAIC=AIC(ADL.1.2.3.a),InfoBIC=BIC(ADL.1.2.3.a)) 
InfoCr <- rbind(InfoCr, newRow)
InfoCr
```



### Also Drop Lag 1 LEI?

```{r, echo=TRUE}
ADL.1.2.3.b <- dynlm(Approving~L(Approving,1) +  
                       DSPIC96_pc_pca +L(DSPIC96_pc_pca, c(1,2))+ 
                       lei_pc  + cei_pc + L(cei_pc, c(1,2)) + 
                       Reagan + BushI + Clinton + BushII + Obama + Trump  +
                       Honey + Inaug +  RAss + IraqKuwait + Desert + 
                       Lehman + Sept11 + Mueller  + Impeach1 + Impeach2 +
                       CovidP + CovidS + ObamaJan + BushJan  + hostages, start=c(1978,4), end=c(2020,12), data=AppEcon.ts)
bgtest(ADL.1.2.3.b, order=12, type="Chisq")
bgtest(ADL.1.2.3.b, order=6, type="Chisq")
bgtest(ADL.1.2.3.b, order=12, type="Chisq")
bgtest(ADL.1.2.3.b, order=6, type="Chisq")

newRow <- data.frame(ModelName='ADL.1.2.3.b',InfoAIC=AIC(ADL.1.2.3.b),InfoBIC=BIC(ADL.1.2.3.b)) 
InfoCr <- rbind(InfoCr, newRow)
InfoCr
```

### Drop LEI entirely?



```{r, echo=TRUE}
ADL.1.2.3.c <- dynlm(Approving~L(Approving,1) +  
                       DSPIC96_pc_pca +L(DSPIC96_pc_pca, c(1,2))+ 
                       cei_pc + L(cei_pc, c(1,2)) + 
                       Reagan + BushI + Clinton + BushII + Obama + Trump +
                       Honey + Inaug +  RAss + IraqKuwait + Desert + 
                       Lehman + Sept11 + Mueller  + Impeach1 + Impeach2 +
                       CovidP + CovidS + ObamaJan + BushJan  + hostages, start=c(1978,4), end=c(2020,12), data=AppEcon.ts)
bgtest(ADL.1.2.3.c, order=12, type="Chisq")
bgtest(ADL.1.2.3.c, order=6, type="Chisq")
bgtest(ADL.1.2.3.c, order=12, type="Chisq")
bgtest(ADL.1.2.3.c, order=6, type="Chisq")

newRow <- data.frame(ModelName='ADL.1.2.3.c',InfoAIC=AIC(ADL.1.2.3.c),InfoBIC=BIC(ADL.1.2.3.c)) 
InfoCr <- rbind(InfoCr, newRow)
InfoCr
```

### Drop lag 2 of CEI from this model?

```{r, echo=TRUE}
ADL.1.2.3.d <- dynlm(Approving~L(Approving,1) +  
                       DSPIC96_pc_pca +L(DSPIC96_pc_pca, c(1,2))+ 
                       cei_pc + L(cei_pc, c(1)) + 
                       Reagan + BushI + Clinton + BushII + Obama + Trump + 
                       Honey + Inaug +  RAss + IraqKuwait + Desert + 
                       Lehman + Sept11 + Mueller  + Impeach1 + Impeach2 + 
                       CovidP + CovidS + ObamaJan + BushJan  + hostages, start=c(1978,4), end=c(2020,12), data=AppEcon.ts)
bgtest(ADL.1.2.3.d, order=12, type="Chisq")
bgtest(ADL.1.2.3.d, order=6, type="Chisq")
bgtest(ADL.1.2.3.d, order=12, type="Chisq")
bgtest(ADL.1.2.3.d, order=6, type="Chisq")

newRow <- data.frame(ModelName='ADL.1.2.3.d',InfoAIC=AIC(ADL.1.2.3.d),InfoBIC=BIC(ADL.1.2.3.d)) 
InfoCr <- rbind(InfoCr, newRow)
InfoCr
```

### Drop lag 0 and 1 but not 2 of CEI?

```{r, echo=TRUE}
ADL.1.2.3.e <- dynlm(Approving~L(Approving,1) +  
                       DSPIC96_pc_pca +L(DSPIC96_pc_pca, c(1,2))+ 
                       L(cei_pc, c(2)) + 
                       Reagan + BushI + Clinton + BushII + Obama + Trump  + 
                       Honey + Inaug +  RAss + IraqKuwait + Desert + 
                       Lehman + Sept11 + Mueller  + Impeach1 + Impeach2 + 
                       CovidP + CovidS + ObamaJan + BushJan  + hostages, start=c(1978,4), end=c(2020,12), data=AppEcon.ts)
bgtest(ADL.1.2.3.e, order=12, type="Chisq")
bgtest(ADL.1.2.3.e, order=6, type="Chisq")
bgtest(ADL.1.2.3.e, order=12, type="Chisq")
bgtest(ADL.1.2.3.e, order=6, type="Chisq")

newRow <- data.frame(ModelName='ADL.1.2.3.e',InfoAIC=AIC(ADL.1.2.3.e),InfoBIC=BIC(ADL.1.2.3.e)) 
InfoCr <- rbind(InfoCr, newRow)
InfoCr
```

### Drop Lag 2 of disposable income growth from model with no LEI and Lag 2 of CEI?

```{r, echo=TRUE}
ADL.1.2.3.f <- dynlm(Approving~L(Approving,1) + 
                       DSPIC96_pc_pca +L(DSPIC96_pc_pca, c(1))+  
                       L(cei_pc, c(2)) + 
                       Reagan + BushI + Clinton + BushII + Obama + Trump  + 
                       Honey + Inaug +  RAss + IraqKuwait + Desert + Lehman + 
                       Sept11 + Mueller  + Impeach1 + Impeach2 + CovidP + 
                       CovidS + ObamaJan + BushJan  + hostages, start=c(1978,4), end=c(2020,12), data=AppEcon.ts)
bgtest(ADL.1.2.3.f, order=12, type="Chisq")
bgtest(ADL.1.2.3.f, order=6, type="Chisq")


newRow <- data.frame(ModelName='ADL.1.2.3.f',InfoAIC=AIC(ADL.1.2.3.f),InfoBIC=BIC(ADL.1.2.3.f)) 
InfoCr <- rbind(InfoCr, newRow)
InfoCr
```

### Drop contemporaneous disposable income growth instead?

```{r, echo=TRUE}
ADL.1.2.3.g <- dynlm(Approving~L(Approving,1) + 
                       L(DSPIC96_pc_pca, c(1,2))+ 
                       L(cei_pc, c(2)) + 
                       Reagan + BushI + Clinton + BushII + Obama + Trump  + 
                       Honey + Inaug +  RAss  + IraqKuwait + Desert + Lehman + 
                       Sept11 + Mueller  + Impeach1 + Impeach2  + CovidP + 
                       CovidS + ObamaJan + BushJan  + hostages, start=c(1978,4), end=c(2020,12), data=AppEcon.ts)
bgtest(ADL.1.2.3.g, order=12, type="Chisq")
bgtest(ADL.1.2.3.g, order=6, type="Chisq")
bgtest(ADL.1.2.3.g, order=12, type="Chisq")
bgtest(ADL.1.2.3.g, order=6, type="Chisq")

newRow <- data.frame(ModelName='ADL.1.2.3.g',InfoAIC=AIC(ADL.1.2.3.g),InfoBIC=BIC(ADL.1.2.3.g)) 
InfoCr <- rbind(InfoCr, newRow)
InfoCr
```

## An automated procedure

The GETS (GEneral To Specific) package contains a function that will take every possible path through the model, testing restrictions and testing for violations of the asymptotic time series assumptions.

To use the function, you must create variables for each lag, rather than use dynlm functions. 

```{r}
# Lag Operator Function
    tslag<-function(y,d=1){
      n<-length(y)
      c(rep(NA,d),y)[1:n]
    }

AppEcon$L1DI <-tslag(AppEcon$DSPIC96_pc_pca, 1)
AppEcon$L2DI <-tslag(AppEcon$DSPIC96_pc_pca, 2)
AppEcon$L1LEI <- tslag(AppEcon$lei_pc, 1)
AppEcon$L2LEI <- tslag(AppEcon$lei_pc, 2)
AppEcon$L1CEI <- tslag(AppEcon$cei_pc, 1)
AppEcon$L2CEI <- tslag(AppEcon$cei_pc, 2)

AppEcon2 <- AppEcon %>%
  select(Approving, lei_pc, L1LEI, L2LEI, cei_pc, L1CEI, L2CEI, DSPIC96_pc_pca, L1DI, L2DI, Reagan, BushI, Clinton, BushII, Obama, Trump, Honey, Inaug, RAss, IraqKuwait, Desert, Lehman, Sept11, Mueller, Impeach1, Impeach2, CovidP, CovidS, ObamaJan, BushJan, hostages, date)

AppEcon2.ts <- ts(AppEcon2, start = c(1978,1), end = c(2020,12), freq = 12)

```

Now we must convert the ts class data to zoo class.

```{r}
AppEcon.zoo <- as.zooreg(AppEcon2.ts)
library(gets)

```


Next, apply the `arx()` function, which models the mean, using the general specification.

The AR-X mean specification can contain an intercept, AR-terms, lagged moving averages of the regressand and other conditioning covariates ('X'). 

arx(y, mc=TRUE, ar=NULL, ewma=NULL, mxreg=NULL, vc=FALSE,
  arch=NULL, asym=NULL, log.ewma=NULL, vxreg=NULL, zero.adj=0.1,
  vc.adj=TRUE, vcov.type=c("ordinary", "white", "newey-west"),
  qstat.options=NULL, normality.JarqueB=FALSE, user.estimator=NULL,
  user.diagnostics=NULL, tol=1e-07, LAPACK=FALSE, singular.ok=TRUE,
  plot=NULL)



```{r}
GUM <- arx(AppEcon.zoo$Approving, ar = 1, mxreg = AppEcon.zoo[,2:32], vcov.type = "white")
```

Finally, apply the `getsm()` function.

The getsm function undertakes multi-path GETS modelling of the mean specification, whereas getsv does the same for the log-variance specification. The diagnostic tests are undertaken on the standardised residuals, and the keep option enables regressors to be excluded from possible removal.

getsm(object, t.pval=0.05, wald.pval=t.pval, vcov.type=NULL, 
    do.pet=TRUE, ar.LjungB=list(lag=NULL, pval=0.025), 
    arch.LjungB=list(lag=NULL, pval=0.025), normality.JarqueB=NULL, 
    user.diagnostics=NULL, info.method=c("sc","aic","aicc", "hq"),
    gof.function=NULL, gof.method=NULL, keep=NULL, include.gum=FALSE,
    include.1cut=TRUE, include.empty=FALSE, max.paths=NULL, tol=1e-07,
    turbo=FALSE, print.searchinfo=TRUE, plot=NULL, alarm=FALSE)


```{r}
getsm(GUM, ar.LjungB=list(lag=12, pval=0.05, info.method="aic",vcov.type = "white", plot=TRUE))
```



# Step Four

```{r, echo=TRUE}
summary(ADL.1.2.3.e)
```

The model selected at this stage excludes the leading economic indicator index entirely and current and lag 1 values of the coincident economic indicator index, but all values of disposable income growth belong in the model. Let's interpret these results. First, let's write out the model (without interventions).

$\hat{A}_t = 3.92 + .0.90A_t + 0.22DI_t + 0.46 DI_{t-1} + 0.37 DI_{t-2} + 102.36 CEI_{t-2}$.


1. Calculate the direct effect of a standard deviation increase in disposable income and in growth in the coincident economic indicator index.
2. Calculate the LRM for for each variable and the standard error.
3. Calculate the Approval equilibrium implied by the model.
4. Plot the IRF and CIRF.
5. Identify the median and mean lag length by printing the standardized CIRF.
6. Calculate the error correction coefficient.
7. Compare the results to the general model.


# What Have We Learned?

**ADL(1,2:3)**


Variable | $\beta_0$ | $\beta_1$ | $\beta_2$ | LRM | LRM se | Median Lag | Mean Lag
--------- | -------- | --------- | --------- | ------ | ------- |----------- | ----------
$DI$ | 0.273 |  0.505 | 0.272 | 11.47 | 3.656 | 8 | 35
$LEI$ | 9.696 | -21.06 | -22.65 | -338.045 | 368.639 | 9| 36
$CEI$ | 23.987 | 40.192 | 121.381 | 1843.91 | 625.03 | 9 | 35
---

Error correction rate: -0.10.

**Final Model**

Variable | $\beta_0$ | $\beta_1$ | $\beta_2$| LRM | LRM se | Median Lag | Mean Lag
--------- | ----------| ---------- | ----- | ------ | ------- |----------- | ----------
$DI$ | 0.205 |  0.463 | 0.367 | 10.108 |  3.087 | 7 | 21
$CEI$ | NA | NA | 102.36 | 998.616 | 229.325 | 7 | 33
---

Error correction rate: -0.10.

