---
title: "Time Series Regression: Part 3 Cointegration"
subtitle: "PLSC 505"
author: "Professor Linn"
date: "March 1, 2022"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '3'
    code_folding: hide
  pdf_document:
    toc: yes
    toc_depth: '3'
  html_notebook:
    number_sections: no
    theme: cerulean
    toc: yes
    toc_depth: 3
  word_document: null
editor_options: 
  markdown: 
    wrap: 72
---

```{=html}
<style type="text/css">

body{ /* Normal  */
      font-size: 26px;
  }
td {  /* Table  */
  font-size: 18px;
}

code.r{ /* Code block */
    font-size: 14px;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 14px;
}
.nobullet li {
  list-style-type: none;
}
}
</style>
```
This help sheet introduces tests for cointegration among I(1) time
series.

# Preliminaries

Load packages and read in the data:

```{r, message=FALSE, warning=FALSE}
library(dynlm)  # To run dynamic LS
suppressPackageStartupMessages(library(stargazer)) # For table output
library(msm) # for delta method
library(lmtest) # for LM tests
library(sandwich) # for various variance-covariance matrices
library(dplyr) # to manipulate data frames
library(ggplot2)
library(rio)
library(aTSA) # for cointegration test function
library(egcm)
library(dplyr)
library(tidyr)
library(forecast) # for checkresiduals function
suppressPackageStartupMessages(source("PGTSadfv2.R")) 
```

```{r}
suppressPackageStartupMessages(library(gdata))
data = read.xls("EconMood_v3.xls")
data$date <- seq(as.Date("1968/4/1"), as.Date("2011/10/30"), by = "quarter")
data <- data %>%
  filter(data<="2010/10/1")
#out1 <- as.Date("1986-01-01")
#data$out1 <- as.numeric(data$date %in% out1)
#out2 <- as.Date("1994-01-01")
#data$out2  <- as.numeric(data$date %in% out2)
#out3 <- as.Date("2004-10-01")
#data$out3  <- as.numeric(data$date %in% out3)

dataQ <- read.csv("QuarterlyApproval.csv")
dataQ$date <- as.Date(dataQ$date)
dataQ <- dataQ %>%
  filter(date >= "1968/4/1" & date <="2010/10/1")

data <- cbind(data, Approving=dataQ$Approving)


```

# Plot the Time Series

```{r, message=FALSE}
df <- data %>%
  select(date, Mood, PolAvg, filtbf) %>%
  gather(key = "series", value = "value", -date)
```

```{r}
ggplot(df, aes(x = date, y = value)) + 
  geom_line(aes(color = series), size = 1) +
   scale_x_date(date_breaks = "24 month", date_labels = "%b\n%Y")+
  scale_color_manual(values = c("#00AFBB", "#E7B800","#E7AFBB", "#00B800"), name = "", 
  labels =  c("Business Expectations", "Public Policy Mood", "Policy Outcomes")) +
  theme_minimal() +
  theme(legend.position="bottom") +
  labs(title="",
       x="Month", 
       y="", 
       subtitle="Fourth Quarter 1968 through Fourth Quarter 2011")
```

# Conduct Unit Root Tests

Use the DF test to determine whether public policy mood (**Mood**),
filetered business expectations (**filtbf**), and average liberalism of
policy outcomes (**PolAvg**) are unit root processes.

## Is Mood a Unit Root Process?

Determine if **Mood** is a unit root process. Assume $D_t=c$.

```{r}
adf.tests(data$Mood, pmax = 8, lmtestorder = 12, type = "constant")
```

```{r}
summary(ur.df(data$Mood, type="drift", lags = 4))
```

## Is PolAvg a Unit Root Process?

```{r}
adf.tests(data$PolAvg, pmax = 12, lmtestorder=12, type="constant")
```

```{r}
adf.tests(data$PolAvg, pmax = 4, lmtestorder=12, type="constant")
```

## Is filtbf a Unit Root Process?

```{r}
adf.tests(data$filtbf, pmax = 24, lmtestorder=12, type="constant")
```

```{r}
adf.tests(data$filtbf, pmax = 23, lmtestorder=12, type="constant")

```

## Conclusions

We cannot reject the null hypothesis that each of these series are unit
root processes, although it's not clear if **Mood** is or is not.

# Engle-Granger Step 1

## Estimate a Cointegrating Regression

```{r}
data.ts <- ts(data, start = c(1968,4), end = c(2010, 3), frequency = 4)
data.ts2 <- as.zooreg(data.ts)
```

We need to decide the appropriate form of deterministics in the test
regression.

-   We want to include a constant to allow there to be a mean difference
    between the series in equilibrium.

-   A trend makes little sense (no series is drifting/trending), so we
    will not include it in the test regression.

-   Previous work suggests mood was impacted by the Vietnam war, the
    Iraq wor, and the Iranian Hostage crisis, so we will allow the mean
    of mood in equilibrium to vary with each intervention:

    -   Values of **Vietnam**: 0, 0.03, 0.06, 0.09, 0.15, 0.24, 0.39,
        0.64, 1, 1, 1, 1, 1, 1 (through 1975), 0, 0, 0 after.
    -   Values of **Iraq**: 0 before first quarter 2003 and 1 after
    -   Values of **IranHost**: 0 before fourth quarter 1979, then 1 for
        6 quarters and then 0.64, 0.39, 0.24, 0.15, 0.09, 0.06, 0.03,
        then 0 after.

```{r}
mystatic <- dynlm(Mood~ PolAvg +  filtbf + IranHost + Vietnam + Iraq, data=data.ts2)
summary(mystatic)
```

## Test Residuals for Cointegration

```{r}
adf.tests(mystatic$residuals, pmax = 12, lmtestorder=12, type="none")
```

```{r}
adf.tests(mystatic$residuals, pmax = 4, lmtestorder=12, type="none")
```

We need to use the tables here:
<https://www.econstor.eu/bitstream/10419/67744/1/616664753.pdf> to
identify the appropriate critical value for the test. Because the static
regression included a constant, we will use Table 2 and consider
$\tau_c$. For the residual-based test we use the case where N=\# of I(1)
variables for which the null of no cointegration is being tested.
Finally, we will use the 0.05 critical value.

To compute the critical value we use the formula:
$\beta_{\infty} + \beta_/T + \beta_/T^2 + \beta_/T^3$, where
$\beta_{\infty}$ is the asymptotic critical value given the form of the
test regression and the remaining $\beta_i$ are used to correct based on
sample size. $T$ is based on the number of observations used for the
residual test regression, here 162. Thus we have:

1.  Find appropriate table based on whether there is a constant or
    constant and trend in the test regression (Here just constant, table
    2).
2.  Find row for N=3. We will use alpha level of .05. Note the various
    $\beta$ values. -3.74066
    , -8.5631, -10.852, 27.982.
3.  Sample size $T$: $T=162$
4.  Calculate following formula for approximate critical value:
    $q(T_i) = \beta_{\infty} + \beta_1(T_i)^{-1} + \beta_2(T_i)^{-2} - \beta_3(T_i)^{-3}$.

$$ -3.74066- 8.5631/162 -10.852/(162^2) + 27.982/(162^3) = -3.793926.
$$

The test statistic value is -4.320. We can reject the null that the
residuals are a unit root process at the 0.05 level of significnce and
thus conclude the series are cointegrated.

## An Alternative Test in R

We can also use the Phillips Peron test with appropriate critical
values. This is the Phillips- Ouliaris cointegration test.

po.test(x, demean = TRUE, lshort = TRUE)

-   x: a matrix or multivariate time series.
-   demean: a logical indicating whether an intercept is included in the
    cointegration regression or not.
-   lshort: a logical indicating whether the short or long version of
    the truncation lag parameter is used.

```{r}
suppressPackageStartupMessages(library(tseries))

myci <- ts(cbind(data$Mood, data$PolAvg, data$filtbf, data$Vietnam, data$Iraq, data$IranHost))
po.test(myci, lshort=TRUE)
```

## Dynamic OLS

Which variables are part of the cointegrating relationship?

```{r}
myDOLS<- dynlm(Mood~  PolAvg +  L(d(PolAvg),-1)+ L(d(PolAvg),-2) +L(d(PolAvg),1) + L(d(PolAvg),2) + filtbf + L(d(filtbf),-1) + L(d(filtbf),-2) + L(d(filtbf),1)  + L(d(filtbf),2) +  IranHost + Vietnam + Iraq, data=data.ts2)

myDOLSHAC <- coeftest(myDOLS, vcov = vcovHAC(myDOLS))
```

```{r}
myDOLS2<- dynlm(Mood~ PolAvg +  filtbf + IranHost + Vietnam + Iraq + L(d(PolAvg),-1) + L(d(PolAvg),1)+ L(d(PolAvg),-2) + L(d(PolAvg),2) + L(d(PolAvg),-3) + L(d(PolAvg),3)+ L(d(PolAvg),-4) + L(d(PolAvg),4) + L(d(filtbf),-1) + L(d(filtbf),1) + L(d(filtbf),-2) + L(d(filtbf),2) + L(d(filtbf),-3) + L(d(filtbf),3) + L(d(filtbf),-4) + L(d(filtbf),4), data=data.ts2)
myDOLSHAC2 <- coeftest(myDOLS2, vcov = vcovHAC(myDOLS2))

stargazer(myDOLSHAC, myDOLSHAC2, type = "text")
```

We can use standard t-tests to test whether the levels variables and
interventions enter the long run relationship. The results suggest
**filtbf** is not part of the cointegerating relationship. Let's
reestimate the static cointegrating regression without this variable.

```{r}
mystatic2 <- dynlm(Mood~ PolAvg + IranHost + Vietnam + Iraq, data=data.ts2)
summary(mystatic2)

```

```{r}
adf.tests(mystatic2$residuals, lmtestorder = 8, pmax=4, type = "none")
```

Test for cointegration:

1.  Find appropriate table based on whether there is a constant or
    constant and trend in the test regression (Here just constant, table
    2).
2.  Find row for N=2. We will use alpha level of .05. Note the various
    $\beta$ values. -3.33613
    , -6.1101, -6.823.
3.  Sample size $T$: $T=162$
4.  Calculate following formula for approximate critical value:
    $q(T_i) = \beta_{\infty} + \beta_1(T_i)^{-1} + \beta_2(T_i)^{-2} - \beta_3(T_i)^{-3}$.

$$ -3.33613 - 6.1101/162 -6.823/(162^2)  = -3.374107.
$$

The test statistic value is -4.166. Reject the null of no cointegration.

We can calculate the disequilibrium from this regression as (or simply
the residuals). I'm calculating them by "hand" here to maintain the
number of observations in the data frame.:

$y_{t-1} - D_{lr} + \lambda_{1} x_{1t} + \lambda_{2} x_{2t} + ...+ \lambda_{n} x_{nt}$

```{r}
data$static.DISEQ <- data$Mood -(mystatic2$coefficients[1] + mystatic2$coefficients[2]*data$PolAvg+ mystatic2$coefficients[3]*data$IranHost + mystatic2$coefficients[4]*data$Vietnam + mystatic2$coefficients[5]*data$Iraq)

data.ts2$static.DISEQ <- mystatic2$residuals

```

We can compare this to the disequilibrium from a respecified DOLS.

```{r}
myDOLS3<- dynlm(Mood~ PolAvg  + IranHost + Vietnam + Iraq + L(d(PolAvg),-1) + L(d(PolAvg),1)+ L(d(PolAvg),-2) + L(d(PolAvg),2) + L(d(PolAvg),-3) + L(d(PolAvg),3)+ L(d(PolAvg),-4) + L(d(PolAvg),4), data=data.ts)
myDOLSHAC3<- coeftest(myDOLS3, vcov = vcovHAC(myDOLS3))
stargazer(myDOLSHAC3, type = "text")
```

$y_{t-1} - D_{lr} + \lambda_{1} x_{1t} + \lambda_{2} x_{2t} + ...+ \lambda_{n} x_{nt}$

```{r}

data$DOLS.DISEQ <- data$Mood -(myDOLS3$coefficients[1] + myDOLS3$coefficients[2]*data$PolAvg+ myDOLS3$coefficients[3]*data$IranHost + myDOLS3$coefficients[4]*data$Vietnam + myDOLS3$coefficients[5]*data$Iraq)

data.ts2$DOLS.DISEQ <- data.ts2$Mood -(myDOLS3$coefficients[1] + myDOLS3$coefficients[2]*data$PolAvg+ myDOLS3$coefficients[3]*data$IranHost + myDOLS3$coefficients[4]*data$Vietnam + myDOLS3$coefficients[5]*data$Iraq)

```

Now let's check the correlation between the two measures of
disequilibrium and plot the disequilibrium.

```{r}
cor(data.ts2$DOLS.DISEQ, data.ts2$static.DISEQ, use = "pairwise.complete.obs")

```

Almost the same!

```{r}
ggplot(data.ts2, aes (x = date, y = static.DISEQ)) + 
  geom_line()+
  geom_line(aes(x = date, y = DOLS.DISEQ), linetype = 2)

```

# Engle-Granger Step 2

We can use the residuals from the static or DOLS regression in an ECM.
To use the disequilibrium, we need to tell R that the data with these
two variables is class ts

```{r}
data.ts <- ts(data, start = c(1968,4), end = c(2010, 3), frequency = 4)
```

```{r}
second.step <- dynlm(d(Mood)~ L(DOLS.DISEQ)  + d(Vietnam) + d(IranHost) + d(Iraq) +  d(PolAvg) + d(filtbf)+  L(d(PolAvg),1) + L(d(filtbf),1) + L(d(Mood),1) + L(d(Mood),4)+ L(d(Mood),8), data=data.ts2)
summary(second.step)
bgtest(second.step, order=4)

checkresiduals(second.step)
```

Here, *because we've determined the variables are cointegrated*, each
righthand-side variable is stationary and we can use standard critical
values for all hypothesis tests.

We'll talk more about interpretation below.

# GECM t-test

To conduct the ECM t-test, we estimate the generalized error correction
model where we include enough lagged changes in the variables to ensure
the residuals are white noise.

```{r}
GECM <- dynlm(d(Mood)~ L(Mood)+ L(PolAvg)+
                Vietnam + Iraq + IranHost+ 
                d(Vietnam) + d(Iraq) + d(IranHost) + 
                d(PolAvg) + L(d(PolAvg), 1) + d(filtbf) +  L(d(filtbf),1) +
                L(d(Mood),1) + L(d(Mood),4) + L(d(Mood),8)
              ,  data=data.ts2)


GECMHAC <- coeftest(GECM, vcov = vcovHAC(GECM))# Heteroscedastic and 

bgtest(GECM, order =4)

checkresiduals(GECM)

```

```{r}
plot(GECM, which=1)
```

```{r}
stargazer(GECMHAC, type="text",
          dep.var.labels = "Changes in Mood",
          report = "vct*",
          column.labels = "GECM")

```

Our interest in this test regression is on the t-statistic on lagged
**Mood**, $\gamma$. Once again we need to identify the appropriate
critical value using the MacKinnon tables (2010). Our GECM does not
contain a trend but does contain a constant, so we will use table 3. We
select the row for k=2 (two I(1) variables in our regression) Finally,
we will use the 0.05 critical value.

1.  Find appropriate table based on whether there is a constant and
    trend (here just constant, table 2).
2.  Find row for k equal to number of independent variables (not
    deterministic). We have 2 will use alpha level of .05. Note the
    various $\theta$ values: -4.7970, -4.7970, -9.04, -5.6, -66.
3.  Compute adjusted sample size $T_a$: $T -(2*k-1) -d$:
    163-(2\*2-1)-4=156
4.  Calculate following formula for approximate critical value:
    $q(T_i) = \theta_{\infty} + \theta_1(T_i)^{-1} + \theta_2(T_i)^{-2} - \theta_3(T_i)^{-3}$.

$$
-4.7970  -9.04/156 -5.6/(156^2) -66/(156^3) = -4.855196
$$

The t-statistic on $\hat{\alpha_1^*}$ in the ecm (the coefficient on
lagged **Mood**, is -5.154, which is further from zero than the critical
value. We can reject the null of no cointegration.

# Interpretation

```{r}
stargazer(second.step, GECMHAC, type = "text",
          title="Cointegration Error Correction Models",
          dep.var.labels="Changes in Mood",
          dep.var.caption = "",
          model.names = FALSE,
          column.labels=c("Two Step", "GECM"))
```

## Long Run Equilibrium

```{r}
LRRGECM <- GECM$coefficients[1]/(-GECM$coefficients[2]) + 
  GECM$coefficients[3]*model.frame(GECM)$"L(PolAvg)"/(-GECM$coefficients[2]) +
  GECM$coefficients[4]*model.frame(GECM)$Vietnam/(-GECM$coefficients[2])+ 
  GECM$coefficients[5]*model.frame(GECM)$Iraq/(-GECM$coefficients[2])+ 
  GECM$coefficients[6]*model.frame(GECM)$IranHost/(-GECM$coefficients[2])



LRRGECM.ts <- ts(LRRGECM, start = c(1971,1), end = c(2010,3), frequency = 4)
LRRStatic.ts <- ts(mystatic2$fitted.values, start = c(1968,4), end = c(2010,3), frequency = 4)
LRRDOLS <- ts(myDOLS3$fitted.values, start = c(1970,1), end = c(2010,3), frequency = 4)

LRE <- ts.union(LRRGECM.ts, LRRStatic.ts, LRRDOLS)
LRE <- na.omit(LRE)
LREdata <-data.frame(LRE=as.matrix(LRE), date=time(LRE))
LRE.longer <- pivot_longer(LREdata, cols= -date, names_to="series", values_to="LRE")
ggplot(LRE.longer, aes(x=date, y = LRE, color = series)) +
  geom_line() 

```

## Impulse Response Functions

```{r}
PolAvg.cirf <- data.frame(period = c(0:20),
                           PolAvgcIRF = NA_real_)

#GECM$coefficients[2] = Lag Mood = -0.3031263
#GECM$coefficients[3] = Lag PolAvg = -0.1165336  * 1 every period after zero
#GECM$coefficients[10] = Delta PolAvg = -0.07515562  * 1 only at zero period
#GECM$coefficients[11] = Lag Delta PolAvg = -0.1455531 * 1 only at first period
#GECM$coefficients[14] = Lag Delta Mood = -0.05470087
#GECM$coefficients[15] = Lag 4 Delta Mood =  0.1617978
#GECM$coefficients[16] =  Lag 8 Delta Mood = 0.06149369 

#Zero period
PolAvg.cirf$PolAvgcIRF[1] <- (1 + GECM$coefficients[2])*0 + GECM$coefficients[3]*0 + GECM$coefficients[10]*1 +  GECM$coefficients[11]*0 + GECM$coefficients[14]*0 + GECM$coefficients[15]*0 + GECM$coefficients[16]*0 

#First period after shock
PolAvg.cirf$PolAvgcIRF[2] <- (1 + GECM$coefficients[2])*PolAvg.cirf$PolAvgcIRF[1] + GECM$coefficients[3]*1 + GECM$coefficients[10]*0 +  GECM$coefficients[11]*1 +
GECM$coefficients[14]*PolAvg.cirf$PolAvgcIRF[1] + GECM$coefficients[15]*0 + GECM$coefficients[16]*0 

#Second period after shock
PolAvg.cirf$PolAvgcIRF[3] <- (1 + GECM$coefficients[2])*PolAvg.cirf$PolAvgcIRF[2] + GECM$coefficients[3]*1 + GECM$coefficients[10]*0 +  GECM$coefficients[11]*0 +
GECM$coefficients[14]*(PolAvg.cirf$PolAvgcIRF[2]-PolAvg.cirf$PolAvgcIRF[1]) + GECM$coefficients[15]*0 + GECM$coefficients[16]*0 

#Third period after shock
PolAvg.cirf$PolAvgcIRF[4] <- (1 + GECM$coefficients[2])*PolAvg.cirf$PolAvgcIRF[3] + GECM$coefficients[3]*1 + GECM$coefficients[10]*0 +  GECM$coefficients[11]*0 +
GECM$coefficients[14]*(PolAvg.cirf$PolAvgcIRF[3]-PolAvg.cirf$PolAvgcIRF[2]) + GECM$coefficients[15]*PolAvg.cirf$PolAvgcIRF[1] + GECM$coefficients[16]*0 

#Fourth period after shock
PolAvg.cirf$PolAvgcIRF[5] <- (1 + GECM$coefficients[2])*PolAvg.cirf$PolAvgcIRF[4] + GECM$coefficients[3]*1 + GECM$coefficients[10]*0 +  GECM$coefficients[11]*0 +
GECM$coefficients[14]*(PolAvg.cirf$PolAvgcIRF[4]-PolAvg.cirf$PolAvgcIRF[3]) + GECM$coefficients[15]*(PolAvg.cirf$PolAvgcIRF[2]-PolAvg.cirf$PolAvgcIRF[1]) + GECM$coefficients[16]*0 

#Fifth period after shock
PolAvg.cirf$PolAvgcIRF[6] <- (1 + GECM$coefficients[2])*PolAvg.cirf$PolAvgcIRF[5] + GECM$coefficients[3]*1 + GECM$coefficients[10]*0 +  GECM$coefficients[11]*0 +
GECM$coefficients[14]*(PolAvg.cirf$PolAvgcIRF[5]-PolAvg.cirf$PolAvgcIRF[4]) + GECM$coefficients[15]*(PolAvg.cirf$PolAvgcIRF[3]-PolAvg.cirf$PolAvgcIRF[2])+ GECM$coefficients[16]*0 

#Sixth period after shock
PolAvg.cirf$PolAvgcIRF[7] <- (1 + GECM$coefficients[2])*PolAvg.cirf$PolAvgcIRF[6] + GECM$coefficients[3]*1 + GECM$coefficients[10]*0 +  GECM$coefficients[11]*0 +
GECM$coefficients[14]*(PolAvg.cirf$PolAvgcIRF[6]-PolAvg.cirf$PolAvgcIRF[5]) + GECM$coefficients[15]*(PolAvg.cirf$PolAvgcIRF[4]-PolAvg.cirf$PolAvgcIRF[3]) + GECM$coefficients[16]*0 

#Seventh period after shock
PolAvg.cirf$PolAvgcIRF[8] <- (1 + GECM$coefficients[2])*PolAvg.cirf$PolAvgcIRF[7] + GECM$coefficients[3]*1 + GECM$coefficients[10]*0 +  GECM$coefficients[11]*0 +
GECM$coefficients[14]*(PolAvg.cirf$PolAvgcIRF[7]-PolAvg.cirf$PolAvgcIRF[6]) + GECM$coefficients[15]*(PolAvg.cirf$PolAvgcIRF[5]-PolAvg.cirf$PolAvgcIRF[4]) + GECM$coefficients[16]*PolAvg.cirf$PolAvgcIRF[1]

#Periods 8+ after shock
for (t in 9:21) {
PolAvg.cirf$PolAvgcIRF[t] <-(1 + GECM$coefficients[2])*PolAvg.cirf$PolAvgcIRF[t-1] + GECM$coefficients[3]*1 + GECM$coefficients[10]*0 +  GECM$coefficients[11]*0 +
GECM$coefficients[14]*(PolAvg.cirf$PolAvgcIRF[t-1]-PolAvg.cirf$PolAvgcIRF[t-2]) + GECM$coefficients[15]*(PolAvg.cirf$PolAvgcIRF[t-4]-PolAvg.cirf$PolAvgcIRF[t-5]) + GECM$coefficients[16]*(PolAvg.cirf$PolAvgcIRF[t-7]-PolAvg.cirf$PolAvgcIRF[t-8])

  }


#### Plot
ggplot(data = PolAvg.cirf, aes(x = period, y = PolAvgcIRF)) +
               geom_bar(stat = "identity") +
               geom_hline(aes(yintercept = 0),
                          colour = "#990000",
                          linetype = "dashed") +
               ylab("CIRF") +
               xlab("Period") +
               ggtitle("CIRF (PolAvg)") +
               theme_bw()

```

```{r}
PolAvg.irf <- data.frame(period = c(0:20),
                           PolAvgIRF = NA_real_)

#Zero period
PolAvg.irf$PolAvgIRF[1] <- (1 + GECM$coefficients[2])*0 + GECM$coefficients[3]*0 + GECM$coefficients[10]*1 +  GECM$coefficients[11]*0 + GECM$coefficients[14]*0 + GECM$coefficients[15]*0 + GECM$coefficients[16]*0 

currentvalue <- PolAvg.irf$PolAvgIRF[1]
#Rest of periods
for (t in 2:21) {
    PolAvg.irf$PolAvgIRF[t] <- PolAvg.cirf$PolAvgcIRF[t] - PolAvg.cirf$PolAvgcIRF[t-1]
    currentvalue <- PolAvg.irf$PolAvgIRF[t]
}
ggplot(data = PolAvg.irf, aes(x = period,y = PolAvgIRF)) +
                geom_bar(stat = "identity") +
                ylab("IRF") +
                xlab("Period") +
                ggtitle("Simple Impulse Response Function PolAvg") +
                theme_bw()
```
